Global patterns of forest autotrophic carbon fluxes

### 2020-04-06 manuscript/manuscript draft.Rmd
updated manuscript draft

The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$AIC) relative to models excluding altitude was generally small (often $\Delta$AIC<2). Hypotheses were accepted if the $\Delta$AIC between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, **unless otherwise specified**. **Becky, are there any instances of this?** Specific analyses are as described below.

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a linear and term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term). If the shape of the relationship [**made biological sense**](https://github.com/forc-db/Global_Productivity/issues/64), and was a significant improvement on the linear relationship ($\Delta$AIC >2), we accepted the polynomial as the best model. We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18). For all FACF, the relationship with latitude was best predicted by the linear model (**but see [this issue]((https://github.com/forc-db/Global_Productivity/issues/64))**).

>>>>

The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$AIC) relative to models excluding altitude was generally small (often $\Delta$AIC<2). Hypotheses were accepted if the $\Delta$AIC between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0. All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. Specific analyses are as described below.

We first examined the relationship between latitude and FACF (Q1; Table 1). We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model. Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term. We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.

@@

All major carbon fluxes decreased linearly with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001). The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18). For all FACF, the relationship with latitude was best predicted by the linear model).
